exploration

### 2021-10-28 code/analysis_code/exploration.Rmd
adding more plots to exploration.Rmd

# load packages and data
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(ggplot2)
library(geodist) #to calculate distance from sample site to each water reclamation facility

@@

# correlation matrices
```{r}
library(GGally) #for produces correlation matrices

#subsetting data to keep only variables we are interested in for the corr plot
mpdatacorr <- mpdata %>% select(-"quarter", 
                                -"month", 
                                -"lat",
                                -"long", 
                                -"year")

@@

>>>>

# load packages and data
```{r}
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse)
library(ggplot2)
library(geodist) #to calculate distance from sample site to each water reclamation facility
library(GGally) #for producing correlation matrices

@@

# correlation matrices
```{r}
#subsetting data to keep only variables we are interested in for the corr plot
mpdatacorr <- mpdata %>% select(-"quarter", 
                                -"month", 
                                -"lat",
                                -"long", 
                                -"year")

@@

# concentration at each site throughout seasons
```{r}
p9 <- mpdata %>% ggplot(aes(x = site, y = particles_l, color = date)) + 
  geom_point() + scale_color_discrete(name = "Date", 
                                     breaks= c("nov_20", "feb_21", "apr_21", "july_21"), 
                                     labels = c("Fall 2020", "Winter 2021", "Spring 2021", "Summer 2021")) +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  labs(title = "Seasonal Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500) #removing outliers to better visualize majority of concentrations

plot(p9)
#save figure
figure_file = here("results","seasonalconc_bysite.png")
ggsave(filename = figure_file, plot=p9)
```

# View concentration by watershed
```{r}
p10 <- mpdata %>% ggplot(aes(x = watershed, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Watershed", x = "Watershed", y = "Particles/Liter") + ylim(0,500)

plot(p10)
#save figure
figure_file = here("results","conc_bywatershed.png")
ggsave(filename = figure_file, plot=p10)
```
# Boxplot for individual sites throughout season
```{r}
p11 <- mpdata %>% ggplot(aes(x = site, y = particles_l, color = )) + 
  geom_boxplot()  +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Microplastic Concentration by Site", x = "Site", y = "Particles/Liter") + ylim(0,500)

plot(p11)
#save figure
figure_file = here("results","conc_bysite.png")
ggsave(filename = figure_file, plot=p11)
```
# concentration over time by watershed
```{r}
# calculate mean concentration in each watershed for each sample date
mean_data <- group_by(mpdata, watershed, date) %>%
             summarise(particles_l = mean(particles_l, na.rm = TRUE))

# create plot using mean data
p12 <- mean_data %>% ggplot(aes(x = date, y = particles_l, color = watershed, group = watershed)) + 
  geom_point() + geom_path() + scale_color_discrete(name = "Watershed")+ 
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "Watershed Microplastic Concentration Over Time", x = "Watershed", y = "Particles/Liter") + scale_x_discrete(labels=c("Fall 2020",
                            "Winter 2021",
                            "Spring 2021",
                            "Summer 2021"))+ 
  ylim(0,300) #removing outliers to better visualize majority of concentrations

plot(p12)
#save figure
figure_file = here("results","watershed_overtime.png")
ggsave(filename = figure_file, plot=p12)
```

Nice! Going forward, I should think about whether to keep or remove the watersheds that were only sampled in the last (July) sampling. I also see that some watershed are missing intermediate data points and are not connected. Would we want to connect and skip over a season, or leave them unconnected (looking at Oconee River)
